\documentclass[12pt]{amsart}

% --------- Encoding & stray Unicode fixes ----------
\DeclareUnicodeCharacter{21D2}{$\Rightarrow$}
\DeclareUnicodeCharacter{2212}{$-$}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

% --------- Core math & graphics ----------
\usepackage{amsmath,amssymb,amsthm}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{tikz}
\usepackage{geometry}
\geometry{margin=1in}

% Figures from both folders (paper figs + R outputs)
\graphicspath{{figs/}{numerical_analysis/sptb_out/}}

% URLs (line-breaking)
\usepackage{url}

% --------- Hyperref LAST ----------
\usepackage[colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue]{hyperref}

% --------- Title & author ----------
\title[A Structural Spline-Penalized Tail Bound]
{A Structural Spline-Penalized Tail Bound for $L$-Functions}

\author{Akbar Akbari Esfahani}
\address{Independent Researcher, Central California Alliance for Health}
\email{akbar.esfahani@gmail.com}

\date{October 2025}

% --------- Theorem Environments ----------
\numberwithin{equation}{section}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{conjecture}[theorem]{Conjecture}

% --------- Page-break policy ----------
% 1) Force every \section to start on a new page.
\makeatletter
\let\SPTB@orig@section\section
\renewcommand{\section}{\clearpage\SPTB@orig@section}
\makeatother
% 2) Helper macro for clear breaks between large content chunks.
\newcommand{\contentbreak}{\clearpage}

\begin{document}

\begin{abstract}
We introduce the Spline–Penalized Tail Bound (SPTB), a finite-window functional that
detects off-critical zeros of automorphic $L$-functions. We prove a rigorous
\emph{detection theorem}: any zero with $\beta>\sigma$ forces exponential growth
$F_\lambda \asymp e^{2(\beta-\sigma)T}$, while the on/left regime obeys an unconditional
polynomial bound $F_\lambda = O(T\log T\log\log T)$. We \emph{conjecture} (and do not prove)
the converse (“Horocycle Conjecture”), so our contribution is a proven detector rather than an equivalence.
We validate numerically (to $0.001\%$) and give a heuristic geometric interpretation.
\end{abstract}

\maketitle
\contentbreak

\tableofcontents
\contentbreak

% --------- Main parts (each begins on a fresh page; sections inside also break) ----------
\input{parts/part1}

\input{parts/part2}

\input{parts/part3}

\input{parts/part4}
\contentbreak

% --------- Appendices ----------
\appendix
\input{appendices/appA}

\input{appendices/appB}

\input{appendices/appC}
\contentbreak

% --------- Bibliography ----------
\begin{thebibliography}{99}

\bibitem{odlyzko} A.~M. Odlyzko, 
\textit{Tables of zeros of the Riemann zeta function},
\url{https://www-odlyzko.dtc.umn.edu/zeta_tables/}.

\bibitem{montgomery-vaughan} H.~L. Montgomery and R.~C. Vaughan,
\textit{The large sieve},
Mathematika \textbf{20} (1973), 119--134.

\bibitem{connes} A. Connes,
\textit{Trace formula in noncommutative geometry and the zeros of the Riemann zeta function},
Selecta Math. (N.S.) \textbf{5} (1999), 29--106.

\bibitem{berry-keating} M.~V. Berry and J.~P. Keating,
\textit{$H = xp$ and the Riemann zeros},
in \textit{Supersymmetry and Trace Formulae: Chaos and Disorder},
Kluwer, 1999, pp.\ 355--367.

\bibitem{balazs-voros} N.~L. Balazs and A. V\"or\"os,
\textit{Chaos on the pseudosphere},
Phys. Rep. \textbf{143} (1986), 109--240.

\bibitem{titchmarsh} E.~C. Titchmarsh,
\textit{The Theory of the Riemann Zeta-Function},
2nd ed., Oxford University Press, 1986.

\end{thebibliography}

\end{document}
